On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-04-27 Manuscript/manuscript.tex
finished first pass

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x^2 = 1$, the variance of $X_t$; and
2) $\phi$, the lag-1 auto-correlation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, \ldots, 1000\}$. In each configuration, we generated 10000
replicates. For each replicate, we constructed three 95\% block bootstrap
CIs for each parameter as described in the last section, and
estimated their actual coverage rates along with their 95\% confidence
intervals from the 10000 replicates. The block bootstrap sampling step was done
with function \texttt{tsboot} from R package \textsl{boot} \citep{boot}, with
block size $\lceil n / l \rceil$. This function by default is an implementation
of moving block bootstrap as described in the previous section.
\jy{size? also it was overlapping blocks. Describe the details here.}

@@

\jy{Make it flow better and conciser}
The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, CI}$ and $\hat\theta_{U, CI}$ be the lower bound and upper bound,
respectively,  for $CI \in \{1, \ldots, n_{CI}\}$, where $n_{CI}$ is the number of 
CIs.
Then $cov = \#\{\hat\theta_{L, CI} < \theta < \hat\theta_{U, CI} \}/n_{CI}$,  
for $CI \in \{1, \ldots, n_{CI}\}$.
The coverage rate is, of course, a
point estimate of a proportion, so a 95\% CI of this coverage
proportion (with $n_{CI} = 10000$) can be constructed. If the proportion
.95 is included in the interval, the block bootstrap method is likely
performing
well. If all values in the interval are below .95, the results would suggest 
that the
method either is providing inaccurate estimation, is underestimating the
process' variability, or a combination of both. If all values in the interval
are above .95, the results suggest that the method is overestimating the
process' variability.

@@

Based on Figure~\ref{fig:mu}, the performance is
about the same for the three interval types studied for $\mu$. As dependence 
gets stronger,
a larger sample size is necessary. For $\mu$, all three types are working asymptotically 
but for finite sizes, they more or less need about the same sample size to have 
the correct coverage. When $\phi < 0$,
$n = 100$ seems to  be large enough to cover $\mu$ at the nominal rate. When
$\phi = 0$, or in other other words, the sample is independently normally
distributed, %from a normal distribution,
$n = 500$ \eds{$n=400$?} seems to be large enough to cover $\mu$ at the nominal 
rate. Samples
from processes with $\phi > 0$ appear to require a larger $n$ than samples from
processes with $\phi < 0$. When $\phi = 0.2$, $n = 300$ seemed to be necessary
for adequate performance, and when $\phi = 0.4$, an even larger $n$ appeared
to be necessary to properly cover $\mu$. It is rare for $\phi$ to be so high,
so in most cases, if $\phi$ is unknown, a sample of $n \geq 500$ should be
used when the target parameter is $\mu$. \eds{I would argue that none of the 
sample sizes considered work well for $\phi=0.2$ or $\phi=0.2$ based on 
Figure~\ref{fig:mu}}

Based on \ref{fig:sigma}, coverage is
notably lower for $BC_a$ intervals, but $\hat{\theta}_{n}$-centered intervals
and basic intervals have similar performance. For these two interval types,
$n = 500$ seems to be necessary to cover $\sigma$ at the nominal rate when
$\phi < 0$. If $\phi = 0$, $n = 400$ seems to large enough to cover $\sigma$ at
the nominal rate. If $\phi = 0.2$, $n = 300$ seems large enough, but if the
process has a greater $\phi$ of $0.4$, $n = 400$ may be necessary. It is more
common for $\phi$ to be $> 0$ in applied analysis, so in most cases, if $\phi$
is unknown, using a sample of $n \geq 400$ should be sufficient for estimation
of $\sigma$. Based on our results, $BC_a$ intervals should not be used to 
estimate $\sigma$.

When block bootstrap is used to estimate $\phi$ using $BC_a$ and percentile
intervals, the assumption that coverage rates asymptotically approach the
nominal rate as $n$ increases appears to be invalid based on \ref{fig:sigma}.
The $BC_a$ interval performs very poorly and deteriorates more noticeable
when the absolute value $\phi$ gets
higher, but even when $\phi = 0$, $\phi$ is still undercovered for $n$ as high
as $700$. Like the $BC_a$ interval, the percentile interval deteriorates in
coverage when $\phi$ is high, although the coverage is not as low.
Interestingly, when $\phi = 0$, the percentile interval appears to actually
over-cover $\phi$. Regardless of whether $\phi$ is positive or negative, these two 
intervals do not perform wellThese results suggest that $BC_a$ and basic percentile
block bootstrap intervals should not be used estimate temporal dependence
factors of time series. On the other hand, $\hat{\theta}_{n}$-centered
intervals seem to recover $\phi$ correctly as long as $n >= 200$, and this
appears to be true regardless of $\phi$, although a greater $\phi$ may result
in a decline in performance.

@@

As shown in the above figures, only $\hat{\theta}_{n}$-centered intervals
managed to solve the deterioration problem with respect to recovering $\phi$,
indicating that this interval should be used as opposed to the uncentered
percentile interval or the more complex $BC_a$ interval when estimating
$\phi$. Increasing $n$ will not fix the performance of either of these
intervals because the assumption that the coverage rate will approach an upper
limit of .95 as $n$ increases is invalid. As expected for
$\hat{\theta}_{n}$-centered intervals, a higher positive $\phi$ demands a
larger $n$ to estimate all parameters, but a negative $\phi$ may require a
lesser $n$. For example, for both $\phi = 0.2$ and $0.4$, $n = 100$ was
sufficient to recover $\mu$ at a rate of approximately 95\%.

@@

Block bootstrap method is a useful method for estimating parameters of a time
series: from simple parameters like the mean to temporal dependence factors.
The motivation for the study is the idea that block bootstrap procedure will
perfectly estimate a parameter of a time series given an infinitely large
sample. The goal for this study was to find what is a good enough finite
sample length for block bootstrap estimation to recover the parameter of a
time series at an acceptable rate.

Again, this relies on the assumption that
there is a size $n$ large enough for the method to work: that is, the method's
performance improves as $n$ increases. Out of the two types of intervals used
in this study, this assumption was found to hold true with respect to
estimating $\phi$ only for $\hat{\theta}_{n}$-centered intervals, whereas both
percentile $BC_a$ intervals exhibited coverage deterioration as $n$
increased.

@@

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods.
This information could prove to be useful for researching using block bootstrap
estimation of time series in domains such as econometrics. Future studies
could investigate if there are types of block bootstrap intervals
constructions that would more appropriately recover the parameters of a time
series. One could also investigate the $n$ needed to make inferences about
other forms of serially dependent data such as an MA process.

>>>>

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x^2 = 1$, the variance of $X_t$; and
2) $\phi$, the lag-1 auto-correlation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, \ldots, 1000\}$. In each configuration, we generated 10000
replicates. For each replicate, we constructed three 95\% block bootstrap
CIs for each parameter as described in the last section, and
estimated their actual coverage rates along with their 95\% confidence
intervals from the 10000 replicates. The block bootstrap sampling step was done
with function \texttt{tsboot} from R package \textsl{boot} \citep{boot}, with
block size $\lceil n / l \rceil$. This function by default is an implementation
of moving block bootstrap as described in the previous section.
\jy{size? also it was overlapping blocks. Describe the details here.}
\eds{Is $l = \lceil n^{1/3} \rceil$? Is it also worth investigating different 
choices of $l$ for fixed (large) $n$?}
\eds{how did you estimate $\mu$, $\sigma_x$ and $\phi$?}

@@

\jy{Make it flow better and conciser}
The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, CI}$ and $\hat\theta_{U, CI}$ be the lower and upper bound,
respectively,  for $CI \in \{1, \ldots, n_{CI}\}$, where $n_{CI}$ is the number 
of CIs.
Then $cov = \#\{\hat\theta_{L, CI} < \theta < \hat\theta_{U, CI} \}/n_{CI}$,  
for $CI \in \{1, \ldots, n_{CI}\}$.
\eds{this notation is a bit sloppy, as it implies $n_{CI}$ changes for each 
value of $CI$; consider using a summation of indicators with a different index}
The coverage rate is, of course, a
point estimate of a proportion, so a 95\% CI of this coverage
proportion (with $n_{CI} = 10000$) can be constructed. If the proportion
0.95 is included in the interval, the block bootstrap method is likely
performing
well. If all values in the interval are below 0.95, the results would suggest 
that the
method either is providing inaccurate estimation, is underestimating the
process' variability, or a combination of both. If all values in the interval
are above .95, the results suggest that the method is overestimating the
process' variability.

@@

Based on Figure~\ref{fig:mu}, the performance is
about the same for the three interval types studied for $\mu$. As dependence 
gets stronger,
a larger sample size is necessary. For $\mu$, all three types are working asymptotically 
but for finite sizes, they more or less need about the same sample size to have 
the correct coverage. When $\phi < 0$,
$n = 100$ seems to  be large enough to cover $\mu$ at the nominal rate. When
$\phi = 0$, or in other other words, the sample is independently normally
distributed, %from a normal distribution,
$n = 500$ \eds{$n=400$?} seems to be large enough to cover $\mu$ at the nominal 
rate. Samples
from processes with $\phi > 0$ appear to require a larger $n$ than samples from
processes with $\phi < 0$. When $\phi = 0.2$, $n = 300$ seemed to be necessary
for adequate performance, and when $\phi = 0.4$, an even larger $n$ appeared
to be necessary to properly cover $\mu$. It is rare for $\phi$ to be so high,
so in most cases, if $\phi$ is unknown, a sample of $n \geq 500$ should be
used when the target parameter is $\mu$. \eds{I would argue that none of the 
sample sizes considered work well for $\phi=0.2$ or $\phi=0.4$ based on 
Figure~\ref{fig:mu} and your criteria above (``If all values in the interval are
 below 0.95, the results would suggest that the
method either is providing inaccurate estimation, is underestimating the
process' variability, or a combination of both.")}

Based on Figure~\ref{fig:sigma}, coverage is
notably lower for $BC_a$ intervals \eds{use consistent notation for this}, 
but $\hat{\theta}_{n}$-centered intervals
and basic intervals have similar performance. For these two interval types,
$n = 500$ seems to be necessary to cover $\sigma$ 
\eds{use $\sigma_x$ instead of $\sigma$} at the nominal rate when
$\phi < 0$. If $\phi = 0$, $n = 400$ seems to large enough to cover $\sigma$ at
the nominal rate. If $\phi = 0.2$, $n = 300$ seems large enough, but if the
process has a greater $\phi$ of $0.4$, $n = 400$ may be necessary. 
\eds{$n=400$ does not appear to be enough to me. I would also argue larger $n$
is needed for the other values of $\phi$ unless the curves asymptote at a value
lower than 0.95. Do the point estimates eventually get closer to the nominal 
level with larger $n$?}  
It is more
common for $\phi$ to be $> 0$ in applied analysis, so in most cases, if $\phi$
is unknown, using a sample of $n \geq 400$ should be sufficient for estimation
of $\sigma$. Based on our results, $BC_a$ intervals should not be used to 
estimate $\sigma$.

When block bootstrap is used to estimate $\phi$ using $BC_a$ and percentile
intervals, the assumption that coverage rates asymptotically approach the
nominal rate as $n$ increases appears to be invalid based on 
Figure~\ref{fig:sigma}.
\eds{Figure~\ref{fig:phi}?}
The $BC_a$ interval performs very poorly and deteriorates more noticeably
when the absolute value $\phi$ gets
higher. Even when $\phi = 0$, $\phi$ is still undercovered for $n$ as high
as $700$. Like the $BC_a$ interval, the percentile interval deteriorates in
coverage when $\phi$ is high, although the coverage is not as low.
Interestingly, when $\phi = 0$, the percentile interval appears to actually
over-cover $\phi$. Regardless of whether $\phi$ is positive or negative, these 
two intervals do not perform well. These results suggest that $BC_a$ and basic 
percentile
block bootstrap intervals should not be used estimate temporal dependence
factors of time series. On the other hand, $\hat{\theta}_{n}$-centered
intervals seem to recover $\phi$ correctly as long as $n \geq 200$, and this
appears to be true regardless of $\phi$, although a greater $\phi$ may result
in a decline in performance.

@@

As shown in the above figures, only $\hat{\theta}_{n}$-centered intervals
managed to solve the deterioration problem with respect to recovering $\phi$,
indicating that this interval should be used as opposed to the uncentered
percentile interval or the more complex $BC_a$ interval when estimating
$\phi$. Increasing $n$ will not fix the performance of either of these
intervals because the assumption that the coverage rate will approach an upper
limit of .95 as $n$ increases is invalid. As expected for
$\hat{\theta}_{n}$-centered intervals, a higher positive $\phi$ demands a
larger $n$ to estimate all parameters, but a negative $\phi$ may require a
lesser $n$. For example, for both $\phi = 0.2$ and $0.4$, $n = 100$ was
sufficient to recover $\mu$ at a rate of approximately 95\%.
\eds{not sure if above paragraph is needed - seems to repeat information in the 
previous two paragraphs}

@@

Block bootstrap is a useful method for estimating parameters of a time
series, from simple parameters like the mean to more complicated temporal
 dependence factors.
%The motivation for the study is the idea that block bootstrap procedure will
%perfectly estimate a parameter of a time series given an infinitely large
%sample. The goal for this study was to find what is a good enough finite
%sample length for block bootstrap estimation to recover the parameter of a
%time series at an acceptable rate.
We know theoretically that the block bootstrap procedure will
perfectly estimate \eds{? interval will correctly cover?  What does the theory 
actually say?} a parameter of a time series given an infinitely large
sample, so the goal for this study was to find the smallest finite
sample length $n$ of a time series in order for the block bootstrap procedure to 
recover its associated parameters at an acceptable rate.

Again, this \eds{what?} relies on the assumption that
there is a size $n$ large enough for the method to work: that is, the method's
performance improves as $n$ increases. Out of the two types of intervals used
in this study, this assumption was found to hold true with respect to
estimating $\phi$ only for $\hat{\theta}_{n}$-centered intervals, whereas both
percentile and $BC_a$ intervals exhibited coverage deterioration as $n$
increased.

@@

This study could be used as a guide for applied statistics courses for students
to generally understand how large of a sample size is sufficient for block
bootstrap to be used versus other inference methods.
This information could also prove to be useful for research using block bootstrap
estimation of time series in domains such as econometrics. Future studies
could investigate if there are types of block bootstrap interval
construction that would more appropriately recover the parameters of a time
series. \eds{are there other methods that you know of?} 
One could also investigate the $n$ needed to make inferences about
other forms of serially dependent data such as a moving average process.
